---
title: "Copetition Homework"
author: "Batuhan Karaköy / Yusuf Yeşilmen / Ayşegül Uzunsoy / Selimcan Biçer - Group 4"
date: "`r Sys.Date()`"
output: slidy_presentation
---

```{r setup, include=FALSE, error=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Import
```{r message=FALSE}
bank_data <- read.csv("bank-full.csv", sep = ';')
dim(bank_data)
```

# Data Cleaning & Transforming
## Checking Column Names
```{r}
colnames(bank_data)

cat("Number of NA's:", sum(is.na(bank_data)), "\n")
```
## Glimpse and Summary
```{r}
library(tidyverse)

glimpse(bank_data)
```

```{r}
summary(bank_data)
```
## Categorical Variables
```{r}
table(bank_data$job)
table(bank_data$marital)
table(bank_data$education)
table(bank_data$default)
table(bank_data$housing)
table(bank_data$loan)
table(bank_data$contact)
table(bank_data$month)
table(bank_data$poutcome)
table(bank_data$y)
```
## Target Variable
```{r}
table(bank_data$y)
prop.table(table(bank_data$y))
```

# Conversion Rates
## Conversion Rate By Job
```{r}
conversion_rate <- prop.table(table(bank_data$y))["yes"]

job_df <- bank_data %>%
  group_by(job) %>%
  summarise(
    total = n(),
    yes_count = sum(y == "yes"),
    response_rate = yes_count / total,
    flag = ifelse(response_rate > conversion_rate, 1, 0)
  )

ggplot(job_df, aes(x = reorder(job, response_rate), y = response_rate, fill = factor(flag))) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = conversion_rate, linetype = "dashed", color = "red") +
  theme_minimal() +
  labs(title = "Response Rate by Job", y = "Response Rate", x = "", fill = "")

```

## Conversion Rate By Marital
```{r}
conversion_rate <- prop.table(table(bank_data$y))["yes"]

marital_df <- bank_data %>%
  group_by(marital) %>%
  summarise(
    total = n(),
    yes_count = sum(y == "yes"),
    response_rate = yes_count / total,
    flag = ifelse(response_rate > conversion_rate, 1, 0)
  )

ggplot(marital_df, aes(x = reorder(marital, response_rate), y = response_rate, fill = factor(flag))) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = conversion_rate, linetype = "dashed", color = "red") +
  theme_minimal() +
  scale_fill_manual(values = c("#888888", "steelblue")) +
  labs(
    title = "Response Rate by Marital Status",
    y = "Response Rate", x = "", fill = ""
  ) +
  theme(legend.position = "none")

```

## Conversion Rate By Education
```{r}
education_df <- bank_data %>%
  group_by(education) %>%
  summarise(
    total = n(),
    yes_count = sum(y == "yes"),
    response_rate = yes_count / total,
    flag = ifelse(response_rate > conversion_rate, 1, 0)
  )

ggplot(education_df, aes(x = reorder(education, response_rate), y = response_rate, fill = factor(flag))) +
  geom_col() +
  coord_flip() +
  geom_hline(yintercept = conversion_rate, linetype = "dashed", color = "red") +
  theme_minimal() +
  scale_fill_manual(values = c("#888888", "steelblue")) +
  labs(
    title = "Response Rate by Education",
    y = "Response Rate", x = "", fill = ""
  ) +
  theme(legend.position = "none")


```

# Visual Summaries
## Histogram: Age
```{r}
ggplot(bank_data, aes(x = age)) +
  geom_histogram(bins = 40, fill = "skyblue", color = "black") +
  theme_minimal() +
  labs(title = "Age Distribution", x = "Age", y = "Count")

```

## Correlation Matrix
```{r}
library(corrplot)

bank_data_corr <- bank_data %>%
  mutate(y = ifelse(y == "yes", 1, 0)) %>%
  select_if(is.numeric)

corrplot(cor(bank_data_corr), method = "color", type = "upper", tl.cex = 0.8)
```

# Linear Models
## Model 1
```{r}
bank_data <- bank_data %>% 
  mutate(y_numeric = ifelse(y == "yes", 1, 0))

lm_model1 <- lm(y_numeric ~ age + duration, data = bank_data)
summary(lm_model1)
```

## Model 2
```{r}
lm_model2 <- lm(y_numeric ~ campaign + pdays + previous, data = bank_data)
summary(lm_model2)
```

## Model 3
```{r}
bank_data <- bank_data %>%
  mutate(across(c(loan, housing, default), as.factor))

lm_model3 <- lm(y_numeric ~ balance + loan + housing + default, data = bank_data)
summary(lm_model3)
```

## Model 4
```{r}
lm_model4 <- lm(y_numeric ~ contact + month + day, data = bank_data)
summary(lm_model4)

```

## R-Square Comperison
```{r}
summary(lm_model1)$r.squared
summary(lm_model2)$r.squared
summary(lm_model3)$r.squared
summary(lm_model4)$r.squared
```

# Logistic Regression
## Model 1
```{r}
library(pscl)

bank_data <- bank_data %>%
  mutate(y = as.factor(ifelse(y == "yes", "Yes", "No")))

logit_model_1 <- glm(y ~ ., data = bank_data, family = "binomial")
summary(logit_model_1)
pR2(logit_model_1)
```

## Model 2
```{r}
logit_model_2 <- glm(y ~ age + duration + campaign + pdays + previous, data = bank_data, family = binomial)

summary(logit_model_2)
pR2(logit_model_2)
```

