---
title: 'Assignment 2: Boston Data'
author: "Batuhan Karaköy / Yusuf Yeşilmen / Selimcan Biçer / Ayşegül Uzunsoy"
date: "`r Sys.Date()`"
output: rmdformats::robobook
---

```{r setup, include=FALSE, error=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective
This project aims to build a logistic regression model to classify crime levels in the Boston dataset. By analyzing socio-economic and housing-related factors, we will predict whether crime is high or low. After performing exploratory data analysis (EDA), we will fit the model and evaluate its performance using a confusion matrix and accuracy score. Finally, we will interpret the results to understand the key factors influencing crime rates.

```{r messages = FALSE}
library(MASS)

data(Boston)

boston <- Boston[,c(2:8,10)]

set.seed(138006) 
thr <- runif(1, min=0.2, max = 0.35)
thr
boston$crimclass <- ifelse(Boston$crim < thr, "Low", "High")
```

# Data Cleaning & Transforming
## Checking Column Names
```{r}
colnames(Boston)
```

## Data Cleaning
```{r}
cat("Number of NA's:", sum(is.na(Boston)), "\n")
```

## Glimpse and Summary
````{r messages = FALSE}
library(tidyverse)
glimpse(Boston)
````

````{r}
summary(Boston)
````
# A Visual Summary: ggpairs
```{r warnings = FALSE}
library(GGally)

ggpairs(boston[, c("rm", "age", "tax", "crimclass")], aes(color = crimclass))
```

```{r}
crime_table <- table(boston$crimclass)
print(crime_table)
```

# crimclass 
```{r}
class(boston$crimclass)
```

```{r}
boston$crimclass <- as.factor(boston$crimclass)

class(boston$crimclass)
```

# Creating the Logistic Regression Model
```{r}
fit.logit <- glm(crimclass ~ ., data = boston, family = binomial)

summary(fit.logit)
```

# Model Evaluation
```{r}
predictions <- predict(logit_model, type = "response")
pred_class <- ifelse(predictions > 0.5, "High", "Low")

conf_matrix <- table(Predicted = pred_class, Actual = boston$crimclass)

accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)

conf_matrix
accuracy
```

# Prediction
```{r}
prd <- predict(fit.logit, type = "response")

prd2 <- ifelse(prd > 0.5, 1, 0)

head(cbind(Actual = boston$crimclass, Predicted = prd2))
```

# Confusion Matrix
```{r}
library(caret)

boston$crimclass <- factor(boston$crimclass, levels = c(0, 1))
prd2 <- factor(prd2, levels = c(0, 1))

confusionMatrix(as.factor(prd2), boston$crimclass)

table(prd2, boston$crimclass)
```

# Conclusion

