---
title: 'Assignment 2: Boston Data'
author: "Batuhan Karaköy / Yusuf Yeşilmen / Selimcan Biçer / Ayşegül Uzunsoy"
date: "`r Sys.Date()`"
output: rmdformats::robobook
---

```{r setup, include=FALSE, error=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Objective
This project aims to build a logistic regression model to classify crime levels in the Boston dataset. By analyzing socio-economic and housing-related factors, we will predict whether crime is high or low. After performing exploratory data analysis (EDA), we will fit the model and evaluate its performance using a confusion matrix and accuracy score. Finally, we will interpret the results to understand the key factors influencing crime rates.

# Data Import
```{r messages = FALSE}
library(MASS)

data(Boston)

boston <- Boston[,c(2:8,10)]

set.seed(138006) 
thr <- runif(1, min=0.2, max = 0.35)
thr
boston$crimclass <- ifelse(Boston$crim < thr, "Low", "High")
```

# Data Cleaning & Transforming
## Checking Column Names
```{r}
colnames(Boston)
```
When we look at the column names of the data, we can see there is no uppercase or blank characters. So we don't have to clean the colnames. 

## Data Cleaning
```{r}
cat("Number of NA's:", sum(is.na(Boston)), "\n")
```
The number of NA's is equal to zero. So we don't have to clean the NA values form data. 

## Glimpse and Summary
````{r messages = FALSE}
library(tidyverse)
glimpse(Boston)
````
From here we see that we have 506 observations. We have 14 columns which all have numberic values.

````{r}
summary(Boston)
````
We see that (`chas`) has so many observations that equal to zero. When we look at the 1st and 3rd quartile values of (`chas`) we can say that at least 75% of the observations are equal to zero. When we look at the median (0) and mean (11.36) values of (`zn`) we see a noticeable difference, with the mean being much higher than the median. This indicates that the distribution is right-skewed. Also the we see that 1st quartile and median values of (`zn`) is equal to zero, which means that 50% of the observations are equal to zero. 

# A Visual Summary: ggpairs
```{r warnings = FALSE}
library(GGally)

ggpairs(boston[, c("rm", "age", "tax", "crimclass")], aes(color = crimclass))
```
From the correlation values, we can see that rm has a negative correlation (-0.240) with age, suggesting that as the average number of rooms increases, the proportion of older buildings decreases. Similarly, tax is negatively correlated (-0.292) with rm, implying that higher property taxes are associated with houses having fewer rooms. On the other hand, (`tax`) shows a moderate positive correlation (0.506) with (`crimclass`), meaning areas with higher taxes tend to have a different crime classification.

```{r}
crime_table <- table(boston$crimclass)
print(crime_table)
```
It shows that there are 237 entries in the "High" crime class and 269 in the "Low" crime class. This indicates a fairly balanced distribution, with slightly more cases in the "Low" category.

# Crimclass Factor Transformation 
```{r}
class(boston$crimclass)
```
First we check the class of (`crimclass`) variable and we see that its a character. So we need to transform it to factor to later use.

```{r}
boston$crimclass <- as.factor(boston$crimclass)

class(boston$crimclass)
```
After the transformation we check the class of (`crimclass`) and now we can see that it is a factor now.

# Creating the Logistic Regression Model
```{r}
fit.logit <- glm(crimclass ~ ., data = boston, family = binomial)

summary(fit.logit)
```
- According to the logistic regression output, several variables significantly affect crime classification. 

- Notably, higher nitrogen oxide concentration (`nox`) is a strong predictor of higher crime levels. 

- Similarly, variables like room count (`rm`), property tax (`tax`), and distance from employment (`dis`) are all positively associated with crime, which may suggest complex socio-economic dynamics.

- Conversely, residential zoning (`zn`) and industrial proportion (`indus`) are negatively associated with crime, indicating that more organized residential or industrial planning might correlate with safer areas.

# Model Evaluation
```{r}
predictions <- predict(fit.logit, type = "response")
pred_class <- ifelse(predictions > 0.5, "High", "Low")

conf_matrix <- table(Predicted = pred_class, Actual = boston$crimclass)

accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)

conf_matrix
cat("Accuracy of the model:", accuracy)
```

# Prediction
```{r}
prd <- predict(fit.logit, type = "response")

prd2 <- ifelse(prd > 0.5, 1, 0)

head(cbind(Actual = boston$crimclass, Predicted = prd2))
```
- The logistic regression model performs very well, achieving an overall accuracy of 88.3%. The confusion matrix shows that the model correctly classified 238 low-crime areas and 209 high-crime areas, with only a small number of misclassifications (29 false negatives and 30 false positives).

- The sensitivity of 88.8% indicates the model is very effective at identifying low-crime areas (class 0), while the specificity of 87.8% shows similar performance for high-crime areas (class 1). The balanced accuracy (88.3%) confirms the model handles class imbalance reasonably well.

- The Kappa statistic of 0.766 suggests substantial agreement between predicted and actual classes, indicating that the model performs significantly better than random guessing.

- Additionally, the p-value from McNemar's test is 1, suggesting there is no significant difference between the false positive and false negative error rates, which indicates a balanced model.

# Confusion Matrix
```{r}
library(caret)

boston$crimclass <- factor(boston$crimclass, levels = c(0, 1))
prd2 <- factor(prd2, levels = c(0, 1))

confusionMatrix(as.factor(prd2), boston$crimclass)

table(prd2, boston$crimclass)
```

# Conclusion
The logistic regression model built on the Boston dataset demonstrates strong predictive performance, achieving an accuracy of over 88%. Key predictors such as nitrogen oxide concentration (`nox`), property tax rates (`tax`), and average number of rooms per dwelling (`rm`) emerged as significant factors positively associated with higher crime levels.

The confusion matrix indicates a balanced performance across both classes, with high sensitivity and specificity. The Kappa statistic further confirms strong agreement between predicted and actual values, suggesting the model is both reliable and well-calibrated.

Overall, the analysis highlights the importance of environmental and economic indicators in predicting urban crime levels. For future work, experimenting with interaction terms or applying non-linear classifiers such as random forests or gradient boosting could further improve predictive power and reveal more complex relationships.

**Note:** Some of the interpretations and explanations were refined with the help of GPT.