---
title: "Assignment 1 Bike Data"
author: "Selimcan Biçer / Batuhan Karaköy / Ayşegül Uzunsoy / Yusuf Yeşilmen - Group 4"
date: "`r Sys.Date()`"
output: 
  cleanrmd::html_document_clean:
    theme: sakura
---

```{r setup, include=FALSE, error=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing Necesarry Libraries
```{r message=FALSE}
library(dplyr)
library(tidyverse)
library(janitor)
library(GGally)
library(corrplot)
library(ggplot2)
```

## Importing dataset
```{r message=FALSE}
bike <- readr::read_csv("bike_data.csv")
```

## Looking for the first and last few observation
```{r}
head(bike)

tail(bike)
```

## First Look: Summary and Glimpse
````{r}
glimpse(bike)

summary(bike)
````
From here we can see that we have three categorical variables which are seasons, holiday and functioning day.

## Looking for the number of NA's
````{r}
sum(is.na(bike))
````
We don't have any NA in our data. So we don't need to clean NA's from data.


# Cleaning & Looking for the structure of the categorical variables
````{r}
bike <- bike |> 
  clean_names()

table(bike$seasons)
table(bike$holiday)
table(bike$functioning_day)
````

## Looking for the structure of the rainfall and snowfall columns
````{r}
table(bike$rainfall)
table(bike$snowfall)
````

## Finding the days with rainfall or snawfall
````{r}
bike_filtered <- bike |>
  filter(rainfall > 0 | snowfall > 0) |> 
  select(rainfall, snowfall)

rain_and_snow_days <- nrow(bike_filtered)
print(paste("Days with rainfall or snowfall:", rain_and_snow_days))
````

## Removing dew point temperature data
````{r}
bike_wo_dew <- bike |> select(-dew_point_temperature)
glimpse(bike_wo_dew)
````

## Creating Some Plots
````{r}
bike |> select(rented_bike_count:temperature) |> 
      ggpairs(upper = list(continous = wrap("cor", family="sans")))
````

````{r}
bike_new <- bike |> select(-humidity, -wind_speed, -visibility, -dew_point_temperature, -solar_radiation)
bike_new |> 
    select_if(is.numeric) |>
    cor () |>
    corrplot( type = "upper",
            insig = "blank",
            diag = FALSE,
            addCoef.col = "grey40")

````

````{r}
bike |> 
  group_by(hour) |>
  summarise(a_rent=mean(rented_bike_count, na.rm = TRUE)) |> 
  ungroup() |> 
  ggplot(aes(x=hour, y=a_rent)) +
  geom_col(just = 0.5, fill = "#c77cff") + 
  labs(
    title = "Average Rented Bike Count",
    x = "Hours of Day",
    y = "Average Rented Bike Count"
  ) + 
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))
````

````{r}
bike |> 
  group_by(temperature) |>
  summarise(a_rent=mean(rented_bike_count, na.rm = TRUE)) |> 
  ungroup() |> 
  ggplot(aes(x=temperature, y=a_rent)) +
  geom_line(just = 0.5, color = "#c77cff") + 
  labs(
    title = "Average Rented Bike Count",
    x = "Temperature",
    y = "Average Rented Bike Count"
  ) + 
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))
````

````{r}
ggplot(bike, aes(x="", y=rented_bike_count, fill=seasons)) +
  geom_bar(stat="identity", width=1) +
    coord_polar("y", start=0) +
  labs(title = "Rented Bike Count by Seasons") +
  theme_void()
````

## Linear Models

````{r}
lm_1 <- lm(rented_bike_count ~ hour + temperature + humidity, data = bike)
summary(lm_1)
````

````{r}
lm_2 <- lm(rented_bike_count ~ seasons + holiday + functioning_day, data = bike)
summary(lm_2)
````

````{r}
lm_3 <- lm(rented_bike_count ~ temperature + wind_speed + solar_radiation, data = bike)
summary(lm_3)
````

````{r}
lm_4 <- lm(rented_bike_count ~ hour, data = bike)
summary(lm_4)
````

