---
title: "MIS 302: Data Analytics II | Assignment 1"
author: "Batuhan Karaköy / Yusuf Yeşilmen / Ayşegül Uzunsoy / Selimcan Biçer / Group 4"
date: "`r Sys.Date()`"
output: rmdformats::robobook
---

```{r setup, include=FALSE, error=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, echo = FALSE}
library(dplyr)
library(tidyverse)
library(janitor)
library(GGally)
library(corrplot)
library(ggplot2)
```

## Objective
The aim of this study is to analyze bike rental data and identify the factors that affect the number of rentals. We start by cleaning and transforming the dataset to ensure accuracy. Then, we perform an exploratory data analysis (EDA) to understand patterns and relationships between variables. Various visualizations help us observe trends, such as the impact of temperature, time of day, and weather conditions on bike rentals. Finally, we build multiple linear regression models to examine how different factors influence rental numbers. This analysis provide insights into bike rental behavior and help predict future trends.

## Data Import
```{r message=FALSE}
bike <- readr::read_csv("bike_data.csv")
```

# Data Cleaning & Transforming
## Checking Column Names
```{r}
colnames(bike)
```
Rented Bike Count, Dew Point Temperature, Solar Radiation, Wind Speed and Functioning Day columns has a blank character and all columns has uppercase characters. We need to transform them into the right format. 

## Data Cleaning
```{r}
bike <- bike |> 
  clean_names()

colnames(bike)

cat("Number of NA's:", sum(is.na(bike)), "\n")
```
Now we don't have any blank or uppercase characters in our column names. Also we don't have any NA's in our data. 

## Glimpse and Summary
````{r}
glimpse(bike)
````
From here we see that we have 8760 observations. We have 14 columns as 4 character and 10 numeric. So, seasons, holiday and functioning_day is our categorical variables. 

````{r}
summary(bike)
````
We see that solar radiation, rainfall and snowfall has so many observations that equal to zero. When we look at the 1st and 3rd quartile values of snowfall and rainfall we can say that at least 75% of the observations are equal to zero. When we look at the median (0.0100) and mean (0.5691) values of solar_radiation we see a noticeable difference, with the mean being much higher than the median. This indicates that the distribution is right-skewed.

## Categorical Variables
````{r}
table(bike$seasons)
table(bike$holiday)
table(bike$functioning_day)
````

## Finding the days with rainfall or snawfall
````{r}
bike_filtered <- bike |>
  filter(rainfall > 0 | snowfall > 0) |> 
  select(rainfall, snowfall)

rain_and_snow_days <- nrow(bike_filtered)
print(paste("Days with rainfall or snowfall:", rain_and_snow_days))
````

## Removing dew point temperature data
````{r}
bike <- bike |>
  select(-dew_point_temperature)

colnames(bike)
````
As we can see dew point temperature is gone.

# Creating Some Visual Summaries
## ggpairs
````{r warnings = FALSE}
bike |> select(rented_bike_count, temperature, humidity) |> 
      ggpairs(upper = list(continous = wrap("cor", family="sans")))
````
Bike rentals increase with warmer temperatures (correlation: 0.539) but slightly decrease as humidity rises (correlation: -0.200). Temperature shows a weak positive link with humidity (correlation: 0.159). Most values for rentals, temperature, and humidity fall within moderate ranges. Warmer weather favors rentals, while high humidity slightly discourages them.

## barplot
````{r}
bike |> 
  group_by(hour) |>
  summarise(a_rent=mean(rented_bike_count, na.rm = TRUE)) |> 
  ungroup() |> 
  ggplot(aes(x=hour, y=a_rent)) +
  geom_col(just = 0.5, fill = "steelblue") + 
  labs(
    title = "Average Rented Bike Count by Hour",
    x = "Hours of Day",
    y = "Average Rented Bike Count"
  ) + 
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))
````
Rentals are high in the early morning, decrease between 3 AM and 6 AM, and rise again after 7 AM, peaking around 5 PM. This suggests that bike rentals are most common during commuting hours, likely when people go to and return from work. Rentals remain relatively high in the evening but drop after 9 PM. Overall, the data shows a clear pattern of higher demand during morning and evening rush hours.

## line plot
````{r warnings = FALSE}
bike |>
  mutate(temp_bin = cut(temperature, breaks = seq(min(temperature), max(temperature), by = 2))) |>
  group_by(temp_bin) %>%
  summarise(a_rent = mean(rented_bike_count, na.rm = TRUE)) |>
  ggplot(aes(x = as.numeric(temp_bin), y = a_rent)) +
  geom_line(size = 0.7, color = "steelblue") +
  xlim(0, 30) +
  labs(
    title = "Average Rented Bike Count by Temperature",
    x = "Temperature",
    y = "Average Rented Bike Count"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"))

````
This graph shows that as the temperature rises, the number of bike rentals increases, reaching a peak at around 20 degrees Celsius. After this point, rentals slightly decline. The graph suggests that moderate temperatures encourage bike rentals, while extreme temperatures might lead to a decrease in demand. This information could help bike rental services plan based on weather conditions.

## stacked bar 
````{r}
bike |>
  mutate(temp_bin = cut(temperature, breaks = seq(floor(min(temperature)), ceiling(max(temperature)), by = 2))) |>
  ggplot(aes(x = temp_bin, y = rented_bike_count, fill = seasons)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Dodge Bar Plot of Bike Rentals by Temperature",
       x = "Temperature",
       y = "Total Rented Bike Count",
       fill = "Seasons") +
  theme_minimal() +
  theme(panel.grid = element_blank(),
    axis.line = element_line(linewidth = 1, colour = "grey80"),
    axis.text.x = element_text(angle = 90, hjust = 1))
````

# Linear Models
## Creating Dummy Variables
```{r}
bike <- bike |>
  mutate(solar_radiation_dummy = ifelse(solar_radiation > 0, 1, 0)) |>
  mutate(rainfall_dummy = ifelse(rainfall > 0, 1, 0)) |>
  mutate(snowfall_dummy = ifelse(snowfall > 0, 1, 0))
```

## Model I
````{r}
lm_1 <- lm(rented_bike_count ~ hour + temperature + humidity + solar_radiation_dummy + snowfall_dummy + rainfall_dummy, data = bike)
summary(lm_1)
````
## Model II
````{r}
lm_2 <- lm(rented_bike_count ~ seasons + holiday + functioning_day, data = bike)
summary(lm_2)
````
## Model III
````{r}
lm_3 <- lm(rented_bike_count ~ temperature + wind_speed + solar_radiation, data = bike)
summary(lm_3)
````
## Model IV
````{r}
lm_4 <- lm(rented_bike_count ~ hour, data = bike)
summary(lm_4)
````